---
title: "Portfolio 1"
author: "Fiona Wang"
date: "2025-02-05"
output: html_document
---

In this Portfolio, we will be creating music together. First I would like to give credit to the following websites:   
https://r-music.rbind.io/posts/2021-04-03-gm/.      

https://stackoverflow.com/questions/52838230/can-you-use-r-to-create-music-note-by-note-such-as-via-midi-or-another-format.   

https://flujoo.github.io/gm/articles/gm.html#timbre


Let's load the package: gm. There are two interpretations of this package: grammar of music and generate music. This is the way to communicate with R about music representation. However, in order to generate the output in the note format as well as to play the audio, you will need to download and install a software called MuseScore. It's easy to download and you will have a lot of fun! Also, if you don't download MuseScore, you will get error messages and won't be able to generate results. 
Download here: https://musescore.org/en/download

We will need to install and load "gm" package.  
```{r load-packages, message=FALSE}
#install and load gm package
#install.packages("gm")
library(gm)
```

Now, let's create a couple notes. 
The meter() function defines the 3/4 time signature, and the line() function sets the musical notes. Here, I am thinking of a waltz type of rhythm, so I set it to 3/4 time signature. Click the play button and listen to this short piece of music. 
```{r pure-notes}
music <-  Music() +
  Meter(3, 4) +
  Line(c("C5", "E5", "E5", "D5", "F5", "F5"))
show(music)
```

It sounds beautiful, but I would like the tempo to be a little slower.
```{r tempo}
music <- music + Tempo(90)
show(music)
```

With only the high notes, I want something low in pitch to accompany that. So, let's add another musical line.
```{r low}
music <- music + 
  (Line(c("C4", "C4", "C4", "D4", "D4", "D4")))
show(music)
```

This is very harmonious! But, let's make things more sophisticated. Wouldn't it be more convenient if we can define each note's duration and create fancy rhythm!
Let's try using pitch and durations.   
```{r pitch_duration}
pd <- Music() +
  Meter(3, 4) +
  Line(c("C5", "E5", "E5", "D5", "F5", "F5")) +
  Line(
    pitches = c("C4", "E4", "D4", "F4"),
    durations = c(2, 1, 2, 1)
  ) +
  Tempo(90)
show(pd)
```

Let's add a pedal to this beautiful piece.
```{r pedal}
pd <- pd + Pedal(1, 4)
show(pd)
```

There is a slight reverb effect, but not much. I will not use pedal anymore.    
I want something other than the sound of Piano. In order to make this easier, I will separately code the two musical lines and then add them together.
```{r timbre}
line1 <- Line(c("C5", "E5", "E5", "D5", "F5", "F5"))
line2 <- Line(
    pitches = c("C4", "E4", "D4", "F4"),
    durations = c(2, 1, 2, 1)
  )
pd <- Music() + Meter(3, 4) + line1 + Instrument(41) + line2 + Instrument(43) + Tempo(100)
show(pd)
```

Beautiful! I always love violin and cello! The loudness of the notes are the same, so it lacks a flow to it. Let's use the dynamic function to change the loudness.
```{r dynamic}
line1 <- Line(c("C5", "E5", "E5", "D5", "F5", "F5"))
line2 <- Line(
    pitches = c("C4", "E4", "D4", "F4"),
    durations = c(2, 1, 2, 1)
  )
pd <- Music() + Meter(3, 4) + 
  line1 + Instrument(41) + Dynamic("mf", 1) + Dynamic("mp", 2) + Dynamic("mp", 3) + 
  line2 + Instrument(43) + Dynamic("mp", 1) + Dynamic("mp", 3) + Tempo(100)
show(pd)
```

As we can see in the output, I have assigned the dynamic markings to the notes. If you like to listen to instrumental music, there are a lot of crescendo and descrescendo in a music piece. Let's add that to the high notes of our piece of music. 
```{r crescendo}
line1 <- Line(c("C5", "E5", "E5", "D5", "F5", "F5"))
line2 <- Line(
    pitches = c("C4", "E4", "D4", "F4"),
    durations = c(2, 1, 2, 1)
  )
pd <- Music() + Meter(3, 4) + 
  line1 + Instrument(41) + Dynamic("pp", 1) + Dynamic("mf", 6) + Hairpin("<", 2, 5) + 
  line2 + Instrument(43) + Dynamic("pp", 1) + Dynamic("pp", 3) + Tempo(100)
show(pd)
```

That's it for today!