---
title: "Portfolio 1"
author: "Fiona Wang"
date: "2025-02-05"
output: html_document
---

In this Portfolio, we will be creating music together. First I would like to give credit to the following websites:   
https://r-music.rbind.io/posts/2021-04-03-gm/
https://stackoverflow.com/questions/52838230/can-you-use-r-to-create-music-note-by-note-such-as-via-midi-or-another-format


Let's load the package: gm. There are two interpretations of this package: grammar of music and generate music. This is the way to communicate with R about music representation. However, in order to generate the output in the note format as well as to play the audio, you will need to download and install a software called MuseScore. It's easy to download and you will have a lot of fun! Also, if you don't download MuseScore, you will get error messages and won't be able to generate results. 
Download here: https://musescore.org/en/download

```{r load-packages, message=FALSE}
#install and load gm package
#install.packages("gm")
library(gm)
```

```{r pure-notes}
music <-   Music() +
  Meter(3, 4) +
  Line(c("C5", "E5", "E5", "D5", "F5", "F5"))
  
show(music)
```
