---
title: "Portfolio 8"
date: 2025-05-04
output: 
  html_document: 
    highlight: kate
    theme: flatly
---

In this portfolio, I will be drawing a graph that captures my favorite 5 restaurants in the U.S.

Let me first give credit to this website: 
https://r-charts.com/spatial/interactive-maps-leaflet/.     

First, let's load libraries
```{r loadlib, message = FALSE}
library(ggplot2)
#install.packages("maps")
library(maps)
#install.packages("leaflet") 
library(leaflet)
```

Then, I will need to create a data set that has the name of the restaurants, the coordinates, and my favorite dish from that restaurant. Favorite dish will pop up when you click on the location.
```{r createdata}
dining <- data.frame(
  name = c("Young Cardinal Cafe and Co","Umiya Raleigh", "Lao Sze Chuan Chinese Restaurant and Bar", "Bostan Uyghur Cuisine", "I'm donut?"),
  lat = c(36.097, 35.819, 40.116, 38.905, 40.757),
  long = c(-80.248, -78.626, -88.243, -77.112, -73.984),
  fav_dish = c("Benedict Egg", "Salmon Sashimi", "Crispy Japanese Tofu with Salted Egg Yolk", "Lagmen", "PB & J")
)
```


Plotting an interacting map
```{r plot}
leaflet(dining) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%  # Clean and modern background
  setView(lng = -85, lat = 38.5, zoom = 4.5) %>%     # Center on the U.S.
  addMarkers(
    lng = ~long,
    lat = ~lat,
    popup = ~paste0(
      name, "<br>",
      "Favorite Dish: ", fav_dish
    )
  )
```

There you have it!